Proposition of a New Experiment to Better Understand the Relation Between Typicality and Prototypes

### 2022-06-22 paper.tex
Added typicality ranking system

			\noindent Since the proposed experiment uses neural networks, there is the need to have a lot of data. In this particular case, there is the need to gather many images. There are some image dataset, intended for computer 
			vision\footnote{\url{https://imerit.net/blog/22-free-image-datasets-for-computer-vision-all-pbm/}}, that could help. Some of the dataset reported there are quite famous, like ImageNet or Google's Open Images. The 
			best for this kind of work, though, is to have a dataset of which every image is part of the same ``macro-category'', like for example a dataset made of images of birds. Even though this is the ideal case,
			most of the dataset that are used for computer vision contain more than one single category. For example, ImageNet, which is one of the most famous image datasets, contains more than 1 million images that are divided 
			into 1000 classes. Generally, it is possible to use these datasets, which have different macro categories by filtering the input images based on the macro category the experiment is based on.

		\subsection{Typicality values\label{sec:typval}}

@@

@@

		\subsection{Test of the second data collection method}

@@

			\noindent The two presented methods proposed for data collection are very different. Both of these have their own peculiarities. The first method can help in having a typicality value for every image collected in the dataset, 
			while the second does not. Given the large number of images that should be used, though, it is not feasible to ask the subjects to rate the typicality value of all the images of the dataset. In fact, it is very likely 
			that a subject that has to rate the typicality of a large number of images will be tired after a certain number of images, polluting the results of the survey. A possible solution to this problem is to 
			divide the survey so that subjects have to rate the typicality of a limited number of images. This method would result in a lower amount of data, but at the same time, a more precise amount of data. Another 
			problem with this method it is that it has a possible bias. In fact, if a person sees a bird, for example a swan, or a duck, and is asked to rate its typicality, most likely will rate it in the higher part 
			of the scale, while if asked to list a certain number of birds it is more likely that it is not present than that it is present.\\
			On the other hand, the second method gives a possibly lower bias, since there are no images that can influence the subject's mind showing images of examples of the category that the subject did not think of. 
			Recalling the previous example with the swan, if a person does not think of the swan, it would not be present in the list created by the subject. This could result in a lower bias, but it comes at the cost 
			of having to interview more people and, most likely, not having enough data for some of the examples. In this approach, the more a bird is typical, the more it appears in the lists and in higher positions. 
			There is an example in the appendix. Even though this method has the advantage of having less bias, has a  great drawback. That is, in particular, the fact that the prototype construction is very difficult.\\
			In order to give the proper weight to both the pros and cons and decide which is the best, the best decision is asking for help to a psychologist or a neuroscientist so that both pros and cons of the two 
			methods have the right weight and it is easier to decide which one is more convenient.
			As alternative, it is possible to try both the methods and see which one gives better results. In this way, both the possibilities will be used and there is the possibility to see which one is better in 
			terms of performance.

@@

			\noindent The second problem is concerning the weight of the image portrays the same example in different positions. Talking about birds, there is an example in Figure \ref{fig:9}. In this case, it is not clear 
			how much should the first image weigh compared to the second one, since they portray the same species of bird in different situations. As said before, there is no algorithm that can average the two images in the 
			way required by this paper, but in this case, most likely, the result would be a mix of two images that would not have much sense, since the swift is in different position, in one it is perched while in the other 
			image is flying. So, to work with this type of approach, an algorithm that computes the average of the two images should be implemented.\\
			If the data are obtained witha the second data collection method proposed, the construction can take two different paths. For the first, which is collecting images based on the survey's answers, the 
			considerations done earlier remain valid, since in the end the feature extraction happens on an image dataset. There is another possibility, though, that is more difficult to implement. In this case, the 
			prototype is built from scratch, using the informations that the subjects gave in the survey. This method theoretically is more accurate than the others, since it can take into account only features 
			that the people answering have in mind. On the other hand, people could not list some features that they consider important because they take them for granted. An example is the presence of the beak in birds. 
			Most likely, the vast majority of people imagine a bird with a beak, but take for granted the presence of the beak, so it is unlikely that it is listed in the features. On the contrary, it is very likely that 
			most people will list, for example, wings in the features that a bird should have, without specifying which type of wings.

@@

			\noindent In case the data collection happened with the second method, the similarity learning happens with a siamese neural network which structure depends on the feature extraction method described. The 
			results, in fact, can be different. If the data collection is made by looking for images of the cited examples, the feature extraction process can follow two routes: the classical one and the machine learning 
			one. In both cases, the similarity learning follows the descriptions in \ref{sec:slcfet} and \ref{sec:slmlfet} respectively, since the resulting prototypes are images. In case the feature extraction did not happen through 
			the collection of images but creating an image following the features listed by the subjects of the survey, the similarity learning still happens with a siamese neural network, but with some differences. In fact 
			the image resulting from this process will have less features than an image retrieved online, because it is created from scratch using the indications of people that took the survey. Involving humans in 
			the feature extraction process, it is possible that the extracted features will not involve colour features, since the colour feature is dependant on the example of the class. If we consider the class ``bird'' as
			example, the colour that people associate with birds is dependent on the species of the considered bird. If the example taken into consideration is the swift, the colour associated with this bird are, most likely, 
			black and white, while for example if the person thinks about a sparrow, the colour thought about are mostly some brown shades. In order to reduce as much as possible colour features, the images fed to 
			the siamese neural network should be preprocessed. Transforming the image to greyscale could be enough, but it should be better to perform a thresholding operation so that the colour features are reduced as much 
			as possible. In both cases, the required siamese network is a network that has two CNNs as embedding networks.

@@

			\noindent The two possible embedding networks, convolutional and fully connected, lead to different internal results but the final goal is the same. In fact, even if with different computations, the siamese 
			network outputs a similarity measure. This output strongly relies on the distance measure. In fact, the same two vectors can have different similarity measures, based on the type of distance that the user 
			decided to implement. For example, suppose that the vectors are $[2,\ 3]$ and $[5, 8]$. If the Manhattan distance is chosen as measure, the distance between the two defined vectors is $8$. 
			Using the euclidean distance instead of the Manhattan distance, the result is $5,83 ca.$. In data science there are some possibilities for distance measurements, a part from euclidean and Manhattan distance. Some 
			examples are the cosine distance or the Hamming distance. Moreover, there could be an ad hoc distance measure, defined for this problem in order to optimize the results obtained by the neural network. 
			This kind of freedom is very beneficial, since it gives the possibility to use different formulas and, seeing the results, there is also the possibility to see which distance works better. This could give an 
			insight on which one of the distance measures is the closest to the one used by human brain in similarity judgement. \\
			Focusing more on the computational aspect of the similarity learning, the more convenient approach is the one involving neural networks feature extraction. In fact, if the features are extracted with 
			neural networks, the siamese network will only need fully connected layers, while it would need convolutional, pooling and fully connected layers if the features are extracted with in a classical way. 
			The only problem about the machine learning feature extraction techniques, as mentioned in \ref{sec:cpc}, is the fact that we are still unable to interpret the features extracted by a CNN.\\
			Talking about the second data collection method, if the chosen method is to ask people to list the characteristics they think are most important, the result will probably involve less features. So, as an 
			educated guess, it is likely that the similarity learning will perform better in case the survey asks to rate the typicality on a scale. In fact, unless the features are too many and the model suffers of the 
			so-called curse of dimensionality, the more the features, the better the performance.

@@

			\noindent One of the first ideas that is an interesting follow-up of this experiment is to compare the results obtained with the two possible feature extraction methods. Even though one of the two methods is not 
			interpretable by humans, it is still interesting trying to understand which of the methods is more efficient in detecting features relevant for the similarity learning. Moreover, if the results 
			indicate that the machine learning techniques are more efficient, could be a push in the research that try to understand the neural networks' feature extraction method. This could be interesting 
			for the future of neural networks and neuroscience. \\
			After the described experiment, there are some natural extensions of the experiment. One of the most natural and obvious extensions of the experiment is to try the same experiment with another category 
			of images. For example, if the first experiment is conducted on a dataset with bird images, it could be interesting to conduct the same experiment on a dataset that uses a different animal like, for 
			example, fishes.\\
			Another experiment that comes as natural is trying to apply some transfer learning techniques. This could be interesting because, as far as we know, the human brain similarity can work 
			in different ways in judging similarity of different images. Applying transfer learning, we could see if the feature extracted for a category are also effective for a different category. This can give 
			extra insight about the typicality and its perception and how it works in human brain. The results can show that human brain innately use the same kind of measure of typicality or a different measure of 
			typicality for different classes. Of course, either result should be investigated more, since a single test is not enough to prove that one of the two possibilities is more likely than the other. Even though 
			it is not enough for a proof, it could be an interesting starting point for neuroscientists to better understand how typicality works in human brain.\\
			Another possibly interesting topic could be examining the results of the survey based on the region of the person answering. In fact, it is possible that different regions, that have slightly different 
			cultures and that may result in slightly different typicality perception. For example, in Europe the eagles are not common, so it is very likely that people will give a high typicality to eagles, while 
			in North America, since the bald eagle is a national symbol, it is easier that people associate with it a higher typicality value. Moreover, there are a lot of different factors that can influence 
			the typicality rates of images in people. For example, a farmer could be more inclined than people that live in a big city to associate a chicken with the concept of ``bird'', and thus giving 
			it a higher typicality.\\
			It is worth to notice that in this paper the prototype is considered as the mean of the features extracted from all the the images of the dataset. A possibility that could be interesting would be 
			interesting to try to experiment the possibility to use different concept of prototype and implementing the correspondent methods of building it.\\
			Lastly, there is another possibility in the medic field. In fact, if this experiment helps in understanding the way that knowledge is organized in the brain, it could be interesting to see if 
			asking people with some knowledge-related or cognitive disorder the results of typicality are similar or not to the ones obtained by asking people without disorders. This experiment could give 
			an insight about how knowledge is organized in brains of people with disorders, and if it is similar to the way brain of people without any disorder organize the knowledge.

>>>>

			\noindent Since the proposed experiment uses neural networks, there is the need to have a lot of data. In this particular case, there is the need to gather many images. There are some image dataset, intended for computer 
			vision\footnote{\url{https://imerit.net/blog/22-free-image-datasets-for-computer-vision-all-pbm/}}, that could help. Some of the dataset reported there are quite famous, like ImageNet or Google's Open Images. The 
			best for this kind of work, though, is to have a dataset of which every image is part of the same ``macro-category'', like for example a dataset made of images of birds. Even though this is the ideal case,
			most of the dataset that are used for computer vision contain more than one single category. For example, ImageNet, which is one of the most famous image datasets, contains more than 1 million images that are divided 
			into 1000 classes. Generally, it is possible to use these datasets, which have different macro categories, by filtering the input images based on the macro category the experiment uses.

		\subsection{Rating based method for typicality values\label{sec:typval}}

@@

		\subsection{Typicality ranking methods for typicality values}

			\noindent Another possibility for collecting typicality values is the typicality ranking. With this method the researcher asks subjects to rank a set of objects based on their typicality. After doing so 
			and collecting the data, the researcher should infer the typicality values based on the ranking. The ranking, as the rating described in \ref{sec:typval}, cannot be performed on all images of the dataset, 
			so it's necessary to split the ranking and rank the images in different forms and then use the informations coming from different subjects and infer the informations about typicality from these answers.

@@

		\subsection{Test of the alternative data collection method}

@@

			\noindent The two presented methods proposed for data collection are very different. Both of these have their own peculiarities. The first method can help in having a typicality value for every image collected in the dataset, 
			while the alternative does not. Given the large number of images that should be used, though, it is not feasible to ask the subjects to rate the typicality value of all the images of the dataset. In fact, it is very likely 
			that a subject that has to rate the typicality of a large number of images will be tired after a certain number of images, polluting the results of the survey. A possible solution to this problem is to 
			divide the survey so that subjects have to rate the typicality of a limited number of images. This method would result in a lower amount of data, but at the same time, a more precise amount of data. Another 
			problem with this method it is that it has a possible bias. In fact, if a person sees a bird, for example a swan, or a duck, and is asked to rate its typicality, most likely will rate it in the higher part 
			of the scale, while if asked to list a certain number of birds it is more likely that it is not present than that it is present.\\
			On the other hand, the second method gives a possibly lower bias, since there are no images that can influence the subject's mind showing images of examples of the category that the subject did not think of. 
			Recalling the previous example with the swan, if a person does not think of the swan, it would not be present in the list created by the subject. This could result in a lower bias, but it comes at the cost 
			of having to interview more people and, most likely, not having enough data for some of the examples. In this approach, the more a bird is typical, the more it appears in the lists and in higher positions. 
			There is an example in the appendix. Even though this method has the advantage of having less bias, has a  great drawback. That is, in particular, the fact that the prototype construction is very difficult.\\
			In order to give the proper weight to both the pros and cons and decide which is the best, the best decision is asking for help to a psychologist or a neuroscientist so that both pros and cons of the two 
			methods have the right weight and it is easier to decide which one is more convenient.
			As alternative, it is possible to try both the methods and see which one gives better results. In this way, both the possibilities will be used and there is the possibility to see which one is better in 
			terms of performance.

@@

			\noindent The second problem is concerning the weight of the image portrays the same example in different positions. Talking about birds, there is an example in Figure \ref{fig:9}. In this case, it is not clear 
			how much should the first image weigh compared to the second one, since they portray the same species of bird in different situations. As said before, there is no algorithm that can average the two images in the 
			way required by this paper, but in this case, most likely, the result would be a mix of two images that would not have much sense, since the swift is in different position, in one it is perched while in the other 
			image is flying. So, to work with this type of approach, an algorithm that computes the average of the two images should be implemented.\\
			If the data are obtained with the alternative data collection method proposed, the construction can take two different paths. For the first, which is collecting images based on the survey's answers, the 
			considerations done earlier remain valid, since in the end the feature extraction happens on an image dataset. There is another possibility, though, that is more difficult to implement. In this case, the 
			prototype is built from scratch, using the informations that the subjects gave in the survey. This method theoretically is more accurate than the others, since it can take into account only features 
			that the people answering have in mind. On the other hand, people could not list some features that they consider important because they take them for granted. An example is the presence of the beak in birds. 
			Most likely, the vast majority of people imagine a bird with a beak, but take for granted the presence of the beak, so it is unlikely that it is listed in the features. On the contrary, it is very likely that 
			most people will list, for example, wings in the features that a bird should have, without specifying which type of wings.

@@

			\noindent In case the data collection happened with the alternative method, the similarity learning happens with a siamese neural network which structure depends on the feature extraction method described. The 
			results, in fact, can be different. If the data collection is made by looking for images of the cited examples, the feature extraction process can follow two routes: the classical one and the machine learning 
			one. In both cases, the similarity learning follows the descriptions in \ref{sec:slcfet} and \ref{sec:slmlfet} respectively, since the resulting prototypes are images. In case the feature extraction did not happen through 
			the collection of images but creating an image following the features listed by the subjects of the survey, the similarity learning still happens with a siamese neural network, but with some differences. In fact 
			the image resulting from this process will have less features than an image retrieved online, because it is created from scratch using the indications of people that took the survey. Involving humans in 
			the feature extraction process, it is possible that the extracted features will not involve colour features, since the colour feature is dependant on the example of the class. If we consider the class ``bird'' as
			example, the colour that people associate with birds is dependent on the species of the considered bird. If the example taken into consideration is the swift, the colour associated with this bird are, most likely, 
			black and white, while for example if the person thinks about a sparrow, the colour thought about are mostly some brown shades. In order to reduce as much as possible colour features, the images fed to 
			the siamese neural network should be preprocessed. Transforming the image to greyscale could be enough, but it should be better to perform a thresholding operation so that the colour features are reduced as much 
			as possible. In both cases, the required siamese network is a network that has two CNNs as embedding networks.

@@

			\noindent The two possible embedding networks, convolutional and fully connected, lead to different internal results but the final goal is the same. In fact, even if with different computations, the siamese 
			network outputs a similarity measure. This output strongly relies on the distance measure. In fact, the same two vectors can have different similarity measures, based on the type of distance that the user 
			decided to implement. For example, suppose that the vectors are $[2,\ 3]$ and $[5, 8]$. If the Manhattan distance is chosen as measure, the distance between the two defined vectors is $8$. 
			Using the euclidean distance instead of the Manhattan distance, the result is $5,83 ca.$. In data science there are some possibilities for distance measurements, a part from euclidean and Manhattan distance. Some 
			examples are the cosine distance or the Hamming distance. Moreover, there could be an ad hoc distance measure, defined for this problem in order to optimize the results obtained by the neural network. 
			This kind of freedom is very beneficial, since it gives the possibility to use different formulas and, seeing the results, there is also the possibility to see which distance works better. This could give an 
			insight on which one of the distance measures is the closest to the one used by human brain in similarity judgement. \\
			Focusing more on the computational aspect of the similarity learning, the more convenient approach is the one involving neural networks feature extraction. In fact, if the features are extracted with 
			neural networks, the siamese network will only need fully connected layers, while it would need convolutional, pooling and fully connected layers if the features are extracted with in a classical way. 
			The only problem about the machine learning feature extraction techniques, as mentioned in \ref{sec:cpc}, is the fact that we are still unable to interpret the features extracted by a CNN.\\
			Talking about the alternative data collection method, if the chosen method is to ask people to list the characteristics they think are most important, the result will probably involve less features. So, as an 
			educated guess, it is likely that the similarity learning will perform better in case the survey asks to rate the typicality on a scale. In fact, unless the features are too many and the model suffers of the 
			so-called curse of dimensionality, the more the features, the better the performance.

@@

			\noindent One of the first ideas that is an interesting follow-up of this experiment is to compare the results obtained with the two possible feature extraction methods. Even though one of the two methods is not 
			interpretable by humans, it is still interesting trying to understand which of the methods is more efficient in detecting features relevant for the similarity learning. Moreover, if the results 
			indicate that the machine learning techniques are more efficient, could be a push in the research that try to understand the neural networks' feature extraction method. This could be interesting 
			for the future of neural networks and neuroscience. \\
			After the described experiment, there are some natural extensions of the experiment. One of the most natural and obvious extensions of the experiment is to try the same experiment with another category 
			of images. For example, if the first experiment is conducted on a dataset with bird images, it could be interesting to conduct the same experiment on a dataset that uses a different animal like, for 
			example, fishes.\\
			Another experiment that comes as natural is trying to apply some transfer learning techniques. This could be interesting because, as far as we know, the human brain similarity can work 
			in different ways in judging similarity of different images. Applying transfer learning, we could see if the feature extracted for a category are also effective for a different category. This can give 
			extra insight about the typicality and its perception and how it works in human brain. The results can show that human brain innately use the same kind of measure of typicality or a different measure of 
			typicality for different classes. Of course, either result should be investigated more, since a single test is not enough to prove that one of the two possibilities is more likely than the other. Even though 
			it is not enough for a proof, it could be an interesting starting point for neuroscientists to better understand how typicality works in human brain.\\
			Another possibly interesting topic could be examining the results of the survey based on the region of the person answering. In fact, it is possible that different regions, that have slightly different 
			cultures and that may result in slightly different typicality perception. For example, in Europe the eagles are not common, so it is very likely that people will give a high typicality to eagles, while 
			in North America, since the bald eagle is a national symbol, it is easier that people associate with it a higher typicality value. Moreover, there are a lot of different factors that can influence 
			the typicality rates of images in people. For example, a farmer could be more inclined than people that live in a big city to associate a chicken with the concept of ``bird'', and thus giving 
			it a higher typicality.\\
			It is worth to notice that in this paper the prototype is considered as the mean of the features extracted from all the the images of the dataset. A possibility that could be interesting would be 
			to try to experiment the possibility to use different concept of prototype and implementing the correspondent methods of building it, possibly even with Smith's concept of prototype.\\
			Lastly, there is another possibility in the medic field. In fact, if this experiment helps in understanding the way that knowledge is organized in the brain, it could be interesting to see if 
			asking people with some knowledge-related or cognitive disorder the results of typicality are similar or not to the ones obtained by asking people without disorders. This experiment could give 
			an insight about how knowledge is organized in brains of people with disorders, and if it is similar to the way brain of people without any disorder organize the knowledge.
